---
title: "Misinformation Spread COVID19"
author: "Sree Kanagala"
output:
  html_document:
    df_print: paged
---
Dataset: [Harvard Dataset](https://doi.org/10.7910/DVN/TIN0SK) <br> <br> <br>

### Data Cleaning 
Load in dataset and create a variable for  
```{r}
library(readxl)
harvard <- read_excel("C:/Users/ksree/Desktop/senior year/comp_app/harvard.xlsx", 
                      sheet = "Sheet16")
View(harvard)

data <- cbind(harvard)
```
<br> <br> <br> Remove binary answers to statements 1-12 for trust and sharing (I use the average per participant for these variables)
```{r}
data = subset(data, select = -c(No, FNtrust1,FNtrust2,FNtrust3,FNtrust4,FNtrust5,FNtrust6,FNtrust7,FNtrust8,
                                FNtrust9,FNtrust10,FNtrust11,FNtrust12,FNtrust13,  FNsharing1,FNsharing2,FNsharing3,FNsharing4,FNsharing5,
                                FNsharing6,FNsharing7,FNsharing8,FNsharing9,FNsharing10,FNsharing11,FNsharing12, FNsharing13))
```

<br> <br> <br>
Profile data to see any immediate anomalies and check variables 
```{r}
View(data)
```

<br> <br> <br>
Needed packages 
```{r}
library(faraway)
library(psych)
```

### Descriptive Statistics 
Get some descriptive statistics for the overall dataset
Identify any visual correlations among variables, visual outliers, mean and skew
```{r}
plot(data)
summary(data)
describe(data)
```

<br> <br> <br>
Create a visual for the correlations existing among variables through a correlation plot
Correlations are not as high as hypothesized 
```{r}
library(GGally)
ggcorr(data, method = c("everything", "pearson")) 
```

### Split Training and Test data
Split data into training and test sets to validate model in the end 
```{r}
require(caTools)
set.seed(4321)
sample = sample.split(data,SplitRatio = 0.75)
train = subset(data,sample ==TRUE) 
test = subset(data, sample==FALSE)

describe(train)
describe(test)
```

### Standardize 
Standardize and scale data so that the mean is 0 and std is 1 (comparing measurements with different units)
Important to scale data after splitting train/test 
```{r}
train$anger_scaled<-scale(train$anger)
train$anxiety_scaled<-scale(train$anxiety)
train$ideology_scaled<-scale(train$ideology)
train$gender_scaled<-scale(train$gender)
train$age_scaled<-scale(train$age)
train$educ_scaled<-scale(train$educ)
train$income_scaled<-scale(train$income)
train$region_scaled<-scale(train$region)
train$st_scaled<-scale(train$st)
train$ss_scaled<-scale(train$ss)

train = subset(train, select = -c(anger, anxiety,ideology, gender, age, educ, income, region, st, ss))
describe(train)

test$anger_scaled<-scale(test$anger)
test$anxiety_scaled<-scale(test$anxiety)
test$ideology_scaled<-scale(test$ideology)
test$gender_scaled<-scale(test$gender)
test$age_scaled<-scale(test$age)
test$educ_scaled<-scale(test$educ)
test$income_scaled<-scale(test$income)
test$region_scaled<-scale(test$region)
test$st_scaled<-scale(test$st)
test$ss_scaled<-scale(test$ss)

test = subset(test, select = -c(anger, anxiety,ideology, gender, age, educ, income, region, st, ss))
describe(test)
```

### PCA
Principle Component Analysis looks to reduce dimensionality through unsupervised learning \  \  
The goal is for the model to explain as much variation as possible by transforming a large set of variables into a smaller one that still contains most of the information in the large set 
```{r}
train.pca <- prcomp(train[,1:10])
train.pca
summary(train.pca)
```

<br> <br> <br>
Visualize how each principle component explains model variation
```{r}
plot(train.pca)
biplot(train.pca)
```

<br> <br> <br>
Reassess correlations with PCA's 
```{r}
library(corrplot)
cor1 <- cor(train.pca$x, method="pearson")
corrplot::corrplot(cor1, method= "color", order = "hclust", tl.pos = 'n')
```

<br> <br> <br>
Save PCA's as a seperate data frame to use for model building 
```{r}
pcs <- as.data.frame(train.pca$x)
data2 <- cbind(train, pcs)
data2 <- subset(data2, select = -c(anger_scaled, anxiety_scaled, ideology_scaled, gender_scaled, age_scaled, educ_scaled,
                                   income_scaled, region_scaled, st_scaled))
data2
```


<br> <br> <br>
Opitmal PCA's is 3   
Run a regression model using the first 3 PCA's 
```{r}
lmodel <- lm(ss_scaled ~ PC1 + PC2 + PC3, data = data2)
summary(lmodel)
```

### Least Squares Model
Compare to a simple least squares regression
```{r}
lmod <- lm(ss_scaled ~ ., data = train)
summary(lmod)
```

<br> <br> <br>
Use forward selection to find most significant features in training dataset
```{r}
library(MASS)
library(olsrr)
model <- lm(ss_scaled ~ ., data = train)
k <- ols_step_forward_p(model)
ols_step_forward_p(model)
plot(k)
```

<br> <br> <br>
Repeat for stepwise selection 
```{r}
k2 <- ols_step_both_p(model)
ols_step_both_p(model)
plot(k2)
summary(k2)
```

<br> <br> <br>
View all possible models 
```{r}
library(olsrr)
k3 <- ols_step_all_possible(model)
k3
plot(k3)
```

<br> <br> <br>
Using these variable selection methods, formally write out each model to compare model accuracy 
```{r}
#pcr
lmodpcr <- lm(train$ss_scaled ~ train.pca$x[,1:3])
#simple 
lmod <- lm(ss_scaled ~ ., data = train)
#forward 
forward <- lm(ss_scaled ~ region_scaled + educ_scaled + ideology_scaled + gender_scaled, data = train)
#stepwise 
stepwise <- lm(ss_scaled ~ st_scaled + region_scaled, data = train)
```

### Cross Validation
Perform cross validation on these 4 models 
```{r}
library(caret)

set.seed(13245)
train.control <- trainControl(method = "repeatedcv", number = 10,repeats=20)
model_leastsquares <- train(ss_scaled ~ ., data = train, method = "lm",
                       trControl = train.control)
model_pcr <- train(ss_scaled ~ PC1 + PC2 + PC3, data = data2, method = "lm",
                   trControl = train.control)
model_forward <- train(ss_scaled ~ region_scaled + educ_scaled + ideology_scaled + gender_scaled, data = train, method = "lm",
                       trControl = train.control)
model_stepwise <- train(ss_scaled ~ st_scaled + region_scaled, data = train, method = "lm",
                        trControl = train.control)

print(model_leastsquares)
print(model_pcr)
print(model_forward)
print(model_stepwise)

```

### Validate 

Set up PCA validation 
```{r}
testpca <- cbind(data2[1:154,])
testpca


PC1 <- train.pca$rotation[,1]
PC2<- train.pca$rotation[,2]
PC3<- train.pca$rotation[,3]
View(testpca)
```

Validate models on test data set
```{r}
model_leastsquares_test <- train(ss_scaled ~ ., data = test, method = "lm", trControl = train.control)
model_PCA_test <- model_pcr_test <- train(ss_scaled ~ PC1 + PC2 + PC3, data = testpca,
                        method = "lm",
                        trControl = train.control)

model_forward_test <- train(ss_scaled ~ region_scaled + educ_scaled + ideology_scaled + gender_scaled, data = test, method =                          "lm", trControl = train.control)

model_stepwise_test <- train(ss_scaled ~ st_scaled + region_scaled, data = test, method = "lm",
                        trControl = train.control)

model_leastsquares_test
model_PCA_test
model_forward_test
model_stepwise_test


```

